Evaluating infectious disease forecasts \\ with allocation scoring rules

### 2024-08-07 alloscore_manuscript/manuscript.Rnw
Made most of the 'minor edits' from Associate Editor

In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to be not
only designed, but also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak
control decisions \citep{marshall2023predictions, bilinski_adaptive_2023}. This contrasts, however, with the
historically standard practice of measuring the quality of disease forecasts using general purpose accuracy and skill
scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For
point forecasts, the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error
(MAE) (e.g., \cite{johansson2016evaluating}) are common choices. For probabilistic forecasts, which are the focus of
this paper, researchers have often relied on the logarithmic score (LS). For example, the LS has been used to evaluate
the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as
forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico \citep{johansson_open_2019}. More
recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts,
the weighted interval score (WIS) \citep{bracher2021evaluating}, have gained prominence. For example, the CRPS was used
to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level in Vietnam
\citep{colon-gonzalez_probabilistic_2021}. And the WIS has been used during the COVID-19 pandemic to evaluate forecasts
of observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal
surveillance systems \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.

@@

Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
impact on and through policy.  And within this body of work we have found the discussion of such a connection to still
be at a formally and quantitatively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
as a way to evaluate the choices made in developing the models.

>>>>

In the wake of the COVID-19 pandemic, this trend has been followed by calls for infectious disease forecasts to be not
only designed, but also evaluated in ways that align specifically with how forecasts can be used to inform such outbreak
control decisions \citep{marshall2023predictions, bilinski_adaptive_2023}. This contrasts, however, with the
historically standard practice of measuring the quality of disease forecasts using general purpose accuracy and skill
scores, especially those that have implementations available in existing software when the relevant outbreak occurs. For
point forecasts, the root mean square error (RMSE) (e.g., \cite{papastefanopoulos2020covid}) and the mean absolute error
(MAE) (e.g., \cite{johansson2016evaluating}) are common choices. For probabilistic forecasts, which are the focus of
this paper, researchers have often relied on the logarithmic score (LS). For example, the LS has been used to evaluate
the skill of US seasonal influenza forecasts \citep{mcgowan_collaborative_2019,reich_collaborative_2019} as well as
forecasts targeting surveillance measures of dengue incidence in Peru and Puerto Rico \citep{johansson_open_2019}. More
recently, the continuous ranked probability score (CRPS) and a discretized version adapted to multi-quantile forecasts,
the weighted interval score (WIS) \citep{bracher2021evaluating}, have gained prominence. For example, the CRPS was used
to assess probabilistic forecasts (based on random effect models) of dengue incidence at the district level in Vietnam
\citep{colon-gonzalez_probabilistic_2021}. Likewise, the WIS has been used during the COVID-19 pandemic to evaluate forecasts
of observed cases, hospitalizations and deaths in the US and Europe, as reported by municipal, state, and federal
surveillance systems \citep{cramer_evaluation_2022,fox_real-time_2022,sherratt2023predictive}.

@@

Despite this now well-developed discussion of the quality-value distinction in the larger forecasting community, we are
aware of only a limited literature attempting to connect the value of \emph{infectious disease} forecasts to their
impact on and through policy.  Even within this body of work we have found the discussion of such a connection to
be at a formally and quantitatively imprecise stage. In \cite{ioannidis2022forecastingCOVIDfailed}, the possible negative
consequences of inaccurate forecasts of infectious disease are discussed, but there is no attempt to quantify the
utility or loss incurred as a result of those forecasts. \cite{bilinski_adaptive_2023} explore ways in which predictive
classifiers of local COVID-19 risk levels in the US could be tuned to policymaker preferences for different costs
associated with over- and under-reaction to disease dynamics, but they do not clearly identify the source of these
costs or how they depend on quantifiable policy choices. A similar discussion related to dengue countermeasures in
Vietnam appears in \cite{colon-gonzalez_probabilistic_2021}.  A novel version of the WIS informally motivated by
utility considerations is developed in \cite{marshall2023predictions}, but the score is not derived in a
decision-theoretic manner.  There is also a thread of literature that frames infectious disease modeling as a component
of a larger system for understanding how policy goals, means, and choices interact and constrain one another. As an
example, \cite{Probert2016decisionMakingFootMouth} explore how policy recommendations ought to flow from a possibly 
incongruous set of simulation-based projection models of a hypothetical foot-and-mouth disease outbreak when there are 
ranges of plausible responses and stakeholder interests. Decision theory plays a prominent role here, but not explicitly
as a way to evaluate the choices made in developing the models.
